---
title: "Demo rownames_to_var"
author: "TheKid"
date: "2023-10-26"
output: html_document
---


```{r setup, include=FALSE} 
# this chunk contains code that sets global options for the entire .Rmd. 
# we use include=FALSE to suppress it from the top of the document, but it will still appear in the appendix. 

knitr::opts_chunk$set(echo = FALSE) # actually set the global chunk options. 
# we set echo=FALSE to suppress code such that it by default does not appear throughout the document. 
# note: this is different from .Rmd default
```

# Introduction  
Below is the function "rownames_to_var". It takes a data frame and converts the data's rownames into a variable.
```{r echo_example, echo=TRUE}
rownames_to_var <- function(data, remove.rownames = TRUE){
  
  names <- rownames(data) #creating vector for all rownames
  
  data_final <- data.frame(names, data) #reposition to ensure names variable is the first column
  
  #if argument remove.rownames is TRUE, rownames will be changed to NULL
  if(remove.rownames == TRUE){
    rownames(data_final) <- NULL
  } else{NULL}
  
  
  return(data_final)
}
```
This function takes in two arguments, data and remove.rownames.   
data: A data frame  
remove.rownames: A logical evaluating to TRUE or FALSE indicating whether rownames in the original data frame should be removed or not. 


# Demonstration
To demonstrate the use of the function I will be using "mtcars": 
```{r}
data("mtcars")

head(mtcars)
```
      
     
The default for the argument remove.rownames has been set to TRUE which tells the function to remove the original rownames in the data frame.
```{r, echo=TRUE}
test_true <- rownames_to_var(mtcars, remove.rownames = TRUE)

head(test_true)
```
Note that a new variable "names" has been added to the data with the original rownames removed and by default R replaces it with a list of integers that corresponds to the row's position.  

On the other hand, if we set remove.rownames to False: 
```{r, echo=TRUE}
test_false <- rownames_to_var(mtcars, remove.rownames = FALSE)

head(test_false)
```
The new variable will still be created with the exception that the original rownames are retained. 


## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```
